Domestic cooking and likelihood

### 2019-11-17 inst/cook.Rmd
work

---
title: "cooking"
output: html_document
---

@@

## R Markdown

Suppose there are five people in a family.  Each day for five days,
one person cooks a meal.  After the fifth day, each person puts the
meals in order of preference, but is not allowed to vote on their own
meal.  This very short vignette shows how to apply generalized
Bradley-Terry to this situation using the `hyper2` package.

```{r cars}
library(hyper2)
H <- hyper2(pnames=paste("p",1:5,sep=""))
H
```

@@

```{r}
H <- H +
    order_likelihood(c(5,2,3,4)) + # person 1 says 5 is the best, 4 the worst.
    order_likelihood(c(1,5,3,4)) + # person 2 says 1 is the best, 4 the worst.
    order_likelihood(c(1,4,2,5)) + # person 3 says 1 is the best, 5 the worst.
    order_likelihood(c(1,2,3,5)) + # person 4 says 1 is the best, 5 the worst.
    order_likelihood(c(2,1,3,4))   # person 5 says 2 is the best, 4 the worst.
H
```

@@

```{r maxlike, echo=FALSE}
dotchart(maxp(H),main="probabilities")
dotchart(log(maxp(H)),main="log probabilities")
```

@@

Now we can assess the hypothesis that person `a` does in fact have the
highest strength of the five.  We can follow the reasoning in the
_icons_ demonstration:

@@

```{r}
library(partitions)
imH <- indep(maxp(H))
f <- function(o){loglik(order_likelihood(o),imH,log=FALSE)}
LL <- apply(perms(5),2,f)
```

Then calculate the pvalue as the probability of obtaining the
observation or an observation more extreme:

@@

>>>>

---
title: "Domestic cooking and likelihood"
output: html_document
---

@@

Suppose there are five people in a family.  Each day for five days,
one person cooks a meal.  After the fifth day, each person considers
the four meals cooked by other family members, and puts them in order
of preference.  This very short vignette shows how to apply
generalized Bradley-Terry to this situation using the `hyper2`
package.

```{r loadlib}
library(hyper2)
H <- hyper2(pnames=paste("p",1:5,sep=""))
H
```

@@

```{r data}
H <- H +
    order_likelihood(c(5,2,3,4)) + # person 1 says 5 is the best, 4 the worst.
    order_likelihood(c(1,5,3,4)) + # person 2 says 1 is the best, 4 the worst.
    order_likelihood(c(1,4,2,5)) + # person 3 says 1 is the best, 5 the worst.
    order_likelihood(c(1,2,3,5)) + # person 4 says 1 is the best, 5 the worst.
    order_likelihood(c(2,1,3,4))   # person 5 says 2 is the best, 4 the worst.
H
```

@@

```{r maxlike, echo=FALSE}
dotchart(maxp(H),pch=16,main="probabilities")
dotchart(log(maxp(H)),pch=16,main="log probabilities")
```

@@

Thus we reject the hypothesis of equal strength.  Now we can assess
the hypothesis that person `a` does in fact have the highest strength
of the five.  We can follow the reasoning in the _icons_
demonstration:

@@

```{r}
library(partitions)
imH <- indep(maxp(H))
f <- function(o){loglik(order_likelihood(o),imH,log=FALSE)}
LL <- apply(perms(5),2,f)
length(LL)
sum(LL)
head(LL)
```

So `LL` is the probability of observing each permutation of 5 objects.
We then calculate the $p$-value as the probability of obtaining the
observation or an observation more extreme; in this case we define a
permutation to be "more extreme" if it has a smaller probability than
the observed permutation of `5,4,3,2,1` (under the null).  This is
easily calculated:

@@

Thus it is reasonable to reject the hypothesis that this particular
observation was drawn from the same distribution as the others, and
infer that it was suspect.
